---
title: "R Data Analysis — Regression, Classification & Football Analytics"
author: "Aziz BENAYED"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width = 7)
```

# Overview

This report brings three strands of R work into one narrative: **linear regression**,
**binary classification with proper evaluation metrics**, and an **exploratory football
analytics** study. Each section states the question, the method, and how to read the result.

---

# 1. Linear regression

## 1a. Marketing spend → sales (`datarium::marketing`)

**Question:** which advertising channel best explains sales?

```{r marketing}
library(datarium)
data("marketing", package = "datarium")
summary(marketing)
model <- lm(sales ~ youtube + facebook + newspaper, data = marketing)
summary(model)
```

Reading it: the coefficients and their p-values rank the channels. In this dataset the
**YouTube** and **Facebook** budgets carry the signal, while **newspaper** typically adds
little once the others are present — a classic multicollinearity / channel-attribution story.

## 1b. Boston housing (`MASS::Boston`)

**Question:** how do neighbourhood attributes relate to median home value (`medv`)?

```{r boston}
library(MASS)
library(corrplot)
corrplot(cor(Boston), method = "color", tl.cex = 0.7)
lm_medv <- lm(medv ~ lstat + age + rm, data = Boston)
summary(lm_medv)
```

The correlation matrix and model show the well-known structure of this dataset: value falls
sharply as **`lstat`** (% lower-status population) rises and grows with **`rm`** (rooms per
dwelling). It's a compact case study in reading coefficients and a correlation heatmap.

---

# 2. Classification — Social Network Ads

**Question:** can we predict whether a user **purchases** from their **age** and
**estimated salary**?

```{r classification}
df <- read.csv("TD1/Social_Network_Ads.csv")
df <- df[, c("Age", "EstimatedSalary", "Purchased")]

set.seed(42)
idx <- sample(seq_len(nrow(df)), 0.75 * nrow(df))
train <- df[idx, ]; test <- df[-idx, ]

model <- glm(Purchased ~ Age + EstimatedSalary, data = train, family = "binomial")
prob <- predict(model, newdata = test, type = "response")
pred <- ifelse(prob > 0.5, 1, 0)
cm <- table(pred, test$Purchased)
cm
```

We evaluate with a hand-written metrics helper (accuracy / specificity / sensitivity) rather
than a black-box call, to make the confusion matrix explicit:

```{r metrics}
metrics <- function(CM) {
  acc <- (CM[1,1] + CM[2,2]) / sum(CM)
  spc <- CM[1,1] / sum(CM[1, ])
  sen <- CM[2,2] / sum(CM[2, ])
  list(accuracy = acc, specificity = spc, sensitivity = sen)
}
metrics(cm)
```

Age and salary are strong, separable predictors of purchase, so a simple logistic model
already classifies well — and the specificity/sensitivity split shows *where* it errs.

---

# 3. Football analytics — Ligue 1 2017/18

**Question:** what team traits go with finishing high in the table?

```{r football}
library(ggplot2)
l1 <- read.csv("TD1/ligue1_17_18.csv", sep = ";")
ggplot(l1, aes(x = ball.possession.per.game.average, y = Points, label = Team)) +
  geom_point(color = "#29B5E8", size = 3) +
  geom_text(vjust = -0.8, size = 3) +
  labs(title = "Points vs average ball possession — Ligue 1 2017/18",
       x = "Avg ball possession per game (%)", y = "Points")
cor(l1$Difference.goals, l1$Points)
```

Unsurprisingly, **goal difference** correlates very strongly with **points**, and the
possession-vs-points scatter separates the dominant sides (PSG) from the pack — a small,
readable sports-analytics EDA.

---

# Takeaways

- Regression: how to fit, read coefficients, and use a correlation heatmap to pick features.
- Classification: a from-scratch confusion-matrix evaluation (accuracy / specificity / sensitivity).
- EDA: framing and visualizing a domain question on real sports data.
